---
title: 'Data Wrangling: Basic Data Manipulation'
author: "Jonathan Freeland"
date: "March 15, 2016"
output:
  html_document:
    theme: null
    css: "style.css"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
library(tidyr)
library(printr)
```

Wrangling performed on a [simple dataset](https://drive.google.com/file/d/0B9vKjeWdQHa5OE5UQ1k4TWJlelU/view) of product orders.
The following code will:

* Clean the 'company' column of misspellings and case differences.
* Separate product code and number.
* Add product categories.
* Add full address for geocoding.
* Create dummy binary variables for company and product category.

***

Read in the data.
```{r, cache=TRUE}
data <- read_csv("refine_original.csv")
```

Clean up the company names.
```{r}
company_names <- c("philips", "akzo", "vanhouten", "unilever")  # acceptable company names
data$company <- lapply(data$company, tolower)                   # all names as lowercase
dist <- adist(data$company, company_names)                      # compute string distances
data$company <- company_names[apply(dist, 1, which.min)]        # choose acceptable name with least distance
```

Separate the product code and number.
```{r}
data <- data %>% separate(`Product code / number`, into = c("product_code", "product_number"), sep = "-")
```

Add product categories.
```{r}
categories <- c("smartphone","tv","laptop","tablet")
names(categories) <- c("p","v","x","q")
data <- data %>% mutate(product_category = categories[product_code])
```

Add full address.
```{r}
data <- data %>% unite(full_address, address, city, country, sep=", ", remove = FALSE)
```

Add dummy columns for company and product category.
```{r}
companies <- (data %>% distinct(company))$company
for(theCompany in companies) # why NROW and not nrow?
  data[paste0("company_",theCompany)] <- as.numeric(data$company == theCompany)

for(theCategory in categories)
  data[paste0("category_",theCategory)] <- as.numeric(data$product_category == theCategory)
```

Export the cleaned data set.
```{r}
write.csv(data, "refine_clean.csv")
```
```{r, echo=FALSE}
head(data)
```